Association between Breast Cancer and Hepatitis C Virus with Age Controlling

### 2022-12-06 manuscript/manuscript.tex
fix discussion

To test the association of breast cancer and Hepatitis C seropositivity, one can use
the chi-square or Fisher's Exact test, depending on the
conditions from the sample size \citep{warner2013testing}. Fisher's
exact test is useful for when the normality assumption is violated 
and the expected values of the 2x2 table are too small. The test uses 
the hypergeometric distribution to test if the probabilities are
the same between the two groups. Thus, we can determine if there is 
more of a risk of breast cancer for those with seropositive Hep C.
In the data section, the data was summarized into 2x2 contingency tables.
The generalzied table is shown in Figure~\ref{fig:table}.
The first step is calculate the probability of the original table occuring,
using the hypergeometric pdf:
\begin{equation}
P_a=[\frac{(a+b+c+d)a!b!c!d!}{(a+c)!(b+d)!(a+b)!(c+d)!}]^-1
\end{equation}

Based on the alternative hypothesis, you then calculate the probability of 
all more extreme tables based on the a value. In this paper, p1 is the
probability of having breast cancer given that the patient is seropositive
and p2 is the probability of having breat cancer given that the patient
is seronegative. Then, we want to determine if there is an association,
specifically if the risk of having breast cancer is higher for those who
are seropositive. The alternative hypothesis is that p1 is greater than p2-
which will be tested in the application section.

@@

After calculating the crude estimate, a Breslow-Day test will test for 
homogeneity of odds ratios between the age stratified data. The generalized
form of stratified 2x2 contingency tables is shown in Figure~\ref{fig:strat}.
The Breslow-Day test statistic will be computed using SAS with the formula:
\begin{equation}
  \chi_{BD}^2=\displaystyle\sum\limits_{i=0}^s\frac{[A_i-E(A_i|\hat{OR}_{M-H})]^2}{Var(A_i|\hat{OR}_{M-H})}
\end{equation}

Based on the results of the Breslow Day test for homogeneity of odds ratios,
the Mantel-Haenszel test will compute a common odds ratio and test if there 
is an association between seropositivity and breast cancer while controlling 
for age. The Mantel-Haenszel test statistic is:
\begin{equation}
  \chi_{M-H}^2=\frac{(|\sum\limits_{i=0}^s A_i - \sum\limits_{i=0}^sE(A_i)|-0.5)^2}
{\sum\limits_{i=0}^s V(A_i)}
\end{equation}

@@

\section{Application}
\label{sec:app}
Fisher's exact test for the pooled data of all patients tested 
\[
H_{0}:p_{1}=p_{2}, H_{a}:p_{1}>p_{2}
\]
at a 0.05 level of confidence.

The test resulted in a p-value of 0.0027, which is less than 
alpha = 0.05. Thus, the null hypothesis is rejected and there 
is evidence suggesting patients who are seropositive have a
higher breast cancer rate than those who are seronegative.

Next, the Breslow Day test checks for homogeneity between the 
stratified ages. 
\[
H_{0}:OR_{1}=OR_{2}, H_{a}:OR_{1}\neq{OR_{2}}
\]
at a 0.01 level of confidence.

The test resulted in a p-value of 0.0133, which is greater than
the alpha value. Thus, we can conclude the strata have similar
odds ratios and proceed with the Mantel-Haenszel procedure.

The Mantel-Haenszel test produced a common odds ratio of 1.6847 and 
a 95 percent confidence interval of (0.9283, 3.0574). The interval
contains 1 so after controlling for age, the association is not supported.

@@

\section{Discussion}
\label{sec:discuss}
  Based on the analysis, it is evident that for women under the age of 45,
seropositivity and breast cancer are strongly associated. As being HCV 
positive can complicate treatment of cancer, this is critical information.
In Egypt, breast cancer remains the most common form or cancer and high 
levels of hepatitus C continue. This study was retrospective so analysis 
was limited to observing odds ratios and the sample size was quite small.
Thus, future studies should aim to be prospective cohort studies so 
researchers can observe how seropisitivity and breast cancer interact 
over time.

Additionally, age is just one factor that can affect the associaiton 
between hepatitus C and breast cancer. Future studies should dive 
deeper into more factors such as weight, reproductive history, 
and alcohol consumption. Addiitonally, future studies can explore how 
seropositivity affects tumor size, disease agression, and more. As 
breast cancer is so prevalent in Egypt, further understanding the 
risk factors is becoming even more important.

>>>>

To test the association of breast cancer and Hepatitis C seropositivity, one can use
the chi-square or Fisher's Exact test, depending on the
conditions from the sample size \citep{warner2013testing}. Fisher's
exact test is useful for when the normality assumption is violated 
and the expected values of the 2x2 table are too small. The test uses 
the hypergeometric distribution to test if the probabilities are
the same between the two groups. Thus, we can determine if there is 
more of a risk of breast cancer for those with seropositive Hep C.

The assumptions of Fisher's Exact Test are that all margins of the table 
are fixed. The total number of subjects with and without the risk factor,
(A + C) and (B + D) must be fixed. The total number of subjects with and 
without the event (A + B) and (C + D) must also be fixed. This is 
equivalent to conditioning on the marginal values.

In the data section, the data was summarized into 2x2 contingency tables.
The generalzied table is shown in Figure~\ref{fig:table}.
The first step is calculate the probability of the original table occuring,
using the hypergeometric probablity mass function:
\begin{equation}
P_a=[\frac{(a+b+c+d)a!b!c!d!}{(a+c)!(b+d)!(a+b)!(c+d)!}]^-1
\end{equation}
The hypergeometric distribution is a discrete distribution, so we use the 
probability mass function for calculations where A is the random variable 
representing the count in the (1,1) cell.

After finding the probability of the original table, you then calculate the probability of 
all more extreme tables based on the a value. Fisher's exact test is often 
a one-sided test based on the research being conducted. In this paper, p1 is the
probability of having breast cancer given that the patient is seropositive
and p2 is the probability of having breat cancer given that the patient
is seronegative. Then, we want to determine if there is an association,
specifically if the risk of having breast cancer is higher for those who
are seropositive. The alternative hypothesis is that p1 is greater than p2-
which will be tested in the application section.

@@

After calculating the crude estimate without age stratfied, 
a Breslow-Day test will test for homogeneity of odds ratios between 
the age stratified data, younger than 45 and age 45 or older. The generalized
form of stratified 2x2 contingency tables is shown in Figure~\ref{fig:strat}.
The risk factor is in the rows while the event of interest is in the columns.
Each strata forms its own table that is used for calculations. 
The Breslow-Day test statistic will be computed using SAS with the formula:
\vspace{0.5cm}
\begin{equation}
  \chi_{BD}^2=\displaystyle\sum\limits_{i=0}^s\frac{[A_i-E(A_i|\hat{OR}_{M-H})]^2}{Var(A_i|\hat{OR}_{M-H})}
\end{equation}
\vspace{0.5cm}

Based on the results of the Breslow Day test for homogeneity of odds ratios,
the Mantel-Haenszel test will compute a common odds ratio and test if there 
is an association between seropositivity and breast cancer while controlling 
for age. The Mantel-Haenszel test can only be conducted if the homogeneity of 
odds ratios is not rejected. The Mantel-Haenszel test statistic is:
\vspace{0.5cm}

\begin{equation}
  \chi_{M-H}^2=\frac{(|\sum\limits_{i=0}^s A_i - \sum\limits_{i=0}^sE(A_i)|-0.5)^2}
{\sum\limits_{i=0}^s V(A_i)}
\end{equation}

\vspace{0.5cm}

@@

This procedure only focuses on odds ratios so it can be used for prospective, 
retrospective, and cross-sectional studies. Additionally, it focuses on the (1,1) 
cell on the table just like Fisher's Exact test. It should be noted that the test 
statistic includes a continuity correction while SAS computes the test statistic 
without a continuity correction.

\section{Application}
\label{sec:app}
As stated in the methods section, we will begin with Fisher's Exact test to 
determine the relationship between HCV seropositivity and breast cancer without 
controlling for age. Then, we will conduct a Breslow-Day test to test for 
homogeneity of odds ratios between the age stratified data. Finally, we will 
perform a Mantel-Haenszel test to compare the strata.

Fisher's exact test for the pooled data of all patients tested 
\[
H_{0}:p_{1}=p_{2}, H_{a}:p_{1}>p_{2}
\]
at a 0.05 level of confidence.

The test resulted in a p-value of 0.0027, which is less than 
alpha = 0.05. Thus, the null hypothesis is rejected and there 
is evidence suggesting patients who are seropositive have a
higher breast cancer rate than those who are seronegative 
without controlling for age.

Next, the Breslow Day test checks for homogeneity between the 
stratified ages. 
\[
H_{0}:OR_{1}=OR_{2}, H_{a}:OR_{1}\neq{OR_{2}}
\]
at a 0.01 level of confidence. Where OR 1 is the odds ratio of 
women who did and did not develop breast cancer with seropositivity 
for women younger than 45 and OR2 is the odds ratio of women who 
did and did not develop breast cancer with seropositivity for women 
45 years and older.

The test resulted in a p-value of 0.0133, which is greater than
the alpha value of 0.05. Thus, we fail to reject the null hypothesis 
and conclude there is not ennough evidence to suggest the strata have different
odds ratios and proceed with the Mantel-Haenszel procedure.

The Mantel-Haenszel test produced a common odds ratio of 1.6847 and 
a 95 percent confidence interval of (0.9283, 3.0574). The interval
contains 1 so after controlling for age, the association is not supported.
Addtionally, we will test the following hypothesis using the Mantel-Haenszel 
test statistic:
\[
H_{0}:OR_{common}=1, H_{a}:OR_{common}\neq{OR_{1}}
\]

The calculated p-value is 0.0773. Therefore, the fail to reject the null 
hypothesis and there is not enough evidence that there is an association 
between HCV seropositivity and breast cancer diagnosis after controlling for 
age.

@@

\section{Discussion}
\label{sec:discuss}
  Based on the analysis, it is evident that for women under the age of 45,
seropositivity and breast cancer are strongly associated. As being HCV 
positive can complicate treatment of cancer, this is critical information.
In Egypt, breast cancer remains the most common form or cancer and high 
levels of hepatitus C continue.

There are several limitations to the study that call into question the 
validity of the results. The small sample size is a result of lack 
of available data. Therefore, it is hard to find a meaningful 
association. The study was retrospective, which limited analysis to 
the odds ratio and not risk difference or relative risk. The method 
of data collection of the control group is also not ideal. Using historical 
data can cause issues as direct populations are not being compared. 
Thus, future studies should aim to be prospective cohort studies so 
researchers can observe how seropositivity and breast cancer interact 
over time.

Additionally, age is just one factor that can affect the associaiton 
between hepatitus C and breast cancer. Future studies should dive 
deeper into more factors such as weight, reproductive history, 
and alcohol consumption. Addiitonally, future studies can explore how 
seropositivity affects tumor size, disease agression, and more. As 
breast cancer is so prevalent in Egypt, further understanding the 
risk factors is becoming even more important. The collaboartion between Egypt 
and Qatar is very promising for emerging research in the health sector. 
With resources and knowledge from Qatar, it is hopeful that Egypt 
can research further into Hepatitis C and breast cancer for the good 
of their population.
